---
title: "Rmd Exercise Template"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

```{r}
gapminder <- read.csv(here::here("data", "gapminder5.csv"))
gapminder$country <-as.character(gapminder$country)
gapminder$continent <-as.character(gapminder$continent)
```

## Print a suitable message if there is any records from 2012
```{r}
check_record_year <-
  function(df, year){
    var <- subset(df, df$year == year)
    if (nrow(var) >0) {cat("Yes, there is any lease one record for year:",year)}
    else {cat("No, there is no record for year:",year)}
  }

check_record_year(gapminder, 2002)
check_record_year(gapminder, 2012)
```


You use the `##` to create a header. Use the header to indicate the question you are answering. After the header, in this space here, annotate your answer, telling us what you did.

```{r chunk name}
# Here is where your R code goes. ``` opens the code chunk, followed by {r}. Then ``` closes the chunk. When you 'knit' the file, the codes will evaluate.

#dat <- read.csv("path/to/data/data.csv")
```

## Including Plots

You can also embed plots, for example:

```{r another chunk name}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
