---
title: "Rmd Exercise Template"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

```{r}
gapminder <- read.csv(here::here("data", "gapminder5.csv"))
gapminder$country <-as.character(gapminder$country)
gapminder$continent <-as.character(gapminder$continent)
```

## Print a suitable message if there is any records from 2012
```{r}
check_record_year <-
  function(df,year){
    var <- df[df$year == year,]
    if (nrow(var) >0) {cat("Yes, there is any lease one record for year:",year)}
    else {cat("No, there is no record for year:",year)}
  }
check_record_year(gapminder, 2002)
check_record_year(gapminder, 2012)
```
##Write a script that finds the mean life expectancy by country for countries whose population is below the mean for the dataset
```{r}
countries <- unique(gapminder$country)
mean_lifeExp <- mean(gapminder$lifeExp)
for (i in countries){
  mean_country <- mean(gapminder$lifeExp[gapminder$country == i])
  if (mean_country < mean_lifeExp)
  {
    cat("\n",i,"'s Mean life expectancy is below mean of the dataset, it is: ",mean_country)
  }
  else 
  {
    cat("\n",i,"'s Mean life expectancy is above mean of the dataset, it is: ",mean_country)
  }
}
```
##write a script that loops through the gapminder data by continent and prints out whether the mean life expectancy is smaller than 50, bertween 50 and 70, or greater than 70
```{r}
continents <- unique(gapminder$continent)
for (i in continents){
  mean_lifeExp <- mean(gapminder$lifeExp[gapminder$continent == i])
  if (mean_lifeExp < 50){cat("\nMean Life Expectancy in",i,"is smaller than 50")}
  else {if (mean_lifeExp>=50 | mean_lifeExp<=70){cat("\nMean Life Expectancy in",i,"is between 50 and 70")}
        else {if (mean_lifeExp>70){cat("\nMean Life Expectancy in",i,"is greater than 70")}
              else {cat("\nMean Life Expectancy in",i,"is outside our expected range")}}

  }
}

```
##Create a function that given a data frame will print the name of each column and the class of data it contains. Use the gapminder dataset. Hint: Use mode() or class() to get the class of the data in each column. Remember that names() or colnames() returns the name of the columns in a dataset.
```{r}
check_col_class<-
  function(df){
    col_names <- colnames(df)
    for (i in col_names){
      class_col <- class(df[df$i,])
      cat("\nThe class of data for colume",i,"is:",class_col)}
  }
```

##Create a function that given a vector will print the mean and the standard deviation of a vector, it will ##optionally also print the median. Hint: include an argument that takes a boolean (TRUE/FALSE) operator and then ##include an if statement.

```{r}
vec_stats<-
  function(vec, include_median = TRUE){
  vec_mean = mean(vec)
  vec_std = sd(vec)
  vec_median =median(vec)
  if(include_median == TRUE){cat("\nThe mean, std and median are:", vec_mean, ",",vec_std,",",vec_median)}
  else{cat("\nThe mean, std are:", vec_mean, ",", vec_std)}
  
  }
```

##Analyzing the relationship

##Use what you’ve learned so far to answer the following questions using the gapminder dataset. Be sure to include some visualizations!

##What is the relationship between GDP per capita and life expectancy? Does this relationship change over time? (Hint: Use the natural log of both variables.)
```{r}
plot(gapminder07$lifeExp ~ gapminder07$gdpPercap, ylab = 'life expenctancy', xlab = 'GDP per capita')
```

##Does the relationship between GDP per capita and life expectacy vary by continent? Make sure you divide the Americas into North and South America.


You use the `##` to create a header. Use the header to indicate the question you are answering. After the header, in this space here, annotate your answer, telling us what you did.

```{r chunk name}
# Here is where your R code goes. ``` opens the code chunk, followed by {r}. Then ``` closes the chunk. When you 'knit' the file, the codes will evaluate.

#dat <- read.csv("path/to/data/data.csv")
```

## Including Plots

You can also embed plots, for example:

```{r another chunk name}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
